---
title: Statistical Rethinking Winter 2020 -- Homework Week 8
author: Alessandro Gentilini \texttt{.@gmail} (just auditing)
date: February 6, 2021
output: pdf_document
---

**1**. Revisit the Reed frog survival data, data(reedfrogs) , and add the predation
and size treatment variables to the varying intercepts model. Consider models with
either predictor alone, both predictors, as well as a model including their interaction.
What do you infer about the causal influence of these predictor variables? Also focus
on the inferred variation across tanks (the $\sigma$ across tanks). Explain why it changes
as it does across models with different predictors included.

***

```{r eval=TRUE, echo=FALSE}
rm(list = ls())
setwd('~/projects/alessandro-gentilini.github.io/bayes/statistical_rethinking_2020/')
```

```{r}
## R code 13.1
library(rethinking)
data(reedfrogs)
d <- reedfrogs
str(d)

## R code 13.2
# make the tank cluster variable
d$tank <- 1:nrow(d)

dat <- list(
    S = d$surv,
    N = d$density,
    tank = d$tank )

    ## R code 13.3
m13.2 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )

precis(m13.2,depth=2)    
plot(precis(m13.2,depth=2))
```

A model with both `pred` and `size`:

```{r}
d$pred_idx <- ifelse(d$pred=='no',1,2)
d$size_idx <- ifelse(d$size=='small',1,2)

dat <- list(
    S = d$surv,
    N = d$density,
    tank = d$tank,
    pred = d$pred_idx,
    my_size = d$size_idx
     )

    ## R code 13.3
m_both <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] + a_size[my_size]+ a_pred[pred],
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_size[my_size] ~ dnorm( a_bar_size , sigma_size ) ,
        a_pred[pred] ~ dnorm( a_bar_pred , sigma_pred ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        a_bar_size ~ dnorm( 0 , 1.5 ) ,
        a_bar_pred ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 ),
        sigma_size ~ dexp( 1 ),
        sigma_pred ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )

precis(m_both,depth=2)    
plot(precis(m_both,depth=2))
```

A model with just `pred`:

```{r}
m_pred <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] + a_pred[pred],
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_pred[pred] ~ dnorm( a_bar_pred , sigma_pred ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        a_bar_pred ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 ),
        sigma_pred ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )

precis(m_pred,depth=2)    
plot(precis(m_pred,depth=2))
```

A model with just `size`:

```{r}
m_size <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] + a_size[my_size],
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_size[my_size] ~ dnorm( a_bar_size , sigma_size ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        a_bar_size ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 ),
        sigma_size ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )

precis(m_size,depth=2)    
plot(precis(m_size,depth=2))
```

I am not able to add an interaction between `pred` and `size`, I had a look at 
page 255 and following but R code 8.24 is about `water` and `shed` being both
ordered variables while in this case `pred` is not ordered while `size` can
be considered ordered.

***

**2**. In 1980, a typical Bengali woman could have 5 or more children in her lifetime.
By the year 2000, a typical Bengali woman had only 2 or 3. You’re going to look at
a historical set of data, when contraception was widely available but many families
chose not to use it. These data reside in data(bangladesh) and come from the
1988 Bangladesh Fertility Survey. Each row is one of 1934 women. There are six
variables, but you can focus on two of them for this practice problem:
(1) district : ID number of administrative district each woman resided in
(2) use.contraception : An indicator (0/1) of whether the woman was using
contraception... 
Now there are 60 values, contiguous integers 1 to 60.
Now, focus on predicting use.contraception , clustered by district_id . Fit
both (1) a traditional fixed-effects model that uses an index variable for district and

(2) a multilevel model with varying intercepts for district. Plot the predicted pro-
portions of women in each district using contraception, for both the fixed-effects
model and the varying-effects model. That is, make a plot in which district ID is on
the horizontal axis and expected proportion using contraception is on the vertical.
Make one plot for each model, or layer them on the same plot, as you prefer. How
do the models disagree? Can you explain the pattern of disagreement? In particu-
lar, can you explain the most extreme cases of disagreement, both why they happen
where they do and why the models reach different inferences?

***

```{r}
data(bangladesh)
d<-bangladesh
d$district_id <- as.integer(as.factor(d$district))
```

```{r}
district_id <- c(unique(d$district_id))
use_c <- c()
district_size <- c()

for( did in district_id ){
    use_c <- c(use_c,sum(d[d$district_id==did,'use.contraception']))
    district_size <- c(district_size,length(d[d$district_id==did,'use.contraception']))
}
```

```{r}
## R code 13.2

dat <- list(
    S = use_c,
    N = district_size,
    district = district_id )

# approximate posterior
m1 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[district] ,
        a[district] ~ dnorm( 0 , 1.5 )
    ), data=dat , chains=4 , log_lik=TRUE )
precis(m1,depth=2)
plot(precis(m1,depth=2))

post1 <- extract.samples(m1)

plot(logistic(apply(post1$a,2, mean)),ylim=c(0,1),main='non varying intercepts, o predicted, + empirical',xlab='district',ylab='proportion using contraception')
points(use_c/district_size,pch='+')
```

```{r}
## R code 13.2


# approximate posterior
m2 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[district] ,
        a[district] ~ dnorm( a_bar , sigma ),
        a_bar ~ dnorm(0,1.5),
        sigma ~ dexp(1)
    ), data=dat , chains=4 , log_lik=TRUE )
precis(m2,depth=2)
plot(precis(m2,depth=2))

post2 <- extract.samples(m2)

plot(logistic(apply(post2$a,2, mean)),ylim=c(0,1),main='varying intercepts, o predicted, + empirical',xlab='district',ylab='proportion using contraception')
points(use_c/district_size,pch='+')
```

```{r}
plot(logistic(apply(post2$a,2, mean)),ylim=c(0,1),main='o varying intercepts, x non varying',xlab='district',ylab='proportion using contraception')
points(logistic(apply(post1$a,2, mean)),pch='x')
```