\input taocpmac

\beginchapter : FORTUNE.

{\quoteformat{Landscape photography is the supreme proof of the photographer, 
and often his supreme disappointment.}
\author ANSEL ADAMS (19??)
}

\medskip

{\quoteformat{I made the arrowheads longer and stronger,
so that they will not disappear so easily on xerox copies.}
\author DONALD E. KNUTH (199?)
}

\medskip

{\quoteformat{I had become thoroughly hooked on the subject
--constantly aware that I was supposed to be doing other things, 
yet unable to resist mathematical beauty.}
\author DONALD E. KNUTH (199?)
}

\medskip

{\quoteformat{Image analysts have a long educational battle to fight 
in the radiological setting in order to bring the message 
that 3D datasets acquired with anisotropies larger than [1 : 2] 
are simply dismissive of the most fundamental concept of 
digital signal processing: The Shannon Sampling Theorem.}
\author ITK (199?)
}

\medskip

{\quoteformat{In fact, a particular pain in the neck
was that usually when they did tomography 
they spaced the slices fairly wide apart, 
so the Z resolution was much lower 
than the X and Y resolutions.}
\author AZRIEL ROSENFELD (1998)
}

\medskip

{\quoteformat{We want to have some theoretical basis for what we're doing. 
We don't want a frivolous or sloppy or second-rate algorithm. 
We don't want some other algorithmist to say, ``You guys are morons''.}
\author ANDREI BRODER (2019)
}

\medskip

{\quoteformat{Computer programming is an art, 
because it applies accumulated knowledge to the world, 
because it requires skill and ingenuity, 
and especially because it produces objects of beauty.}
\author DONALD E KNUTH (1974)
}

\medskip

{\quoteformat{Thus, for example, Donald Knuth took Chomsky's book
on structural linguistics with him on his honeymoon. 
In spite of this Don is still married to Jill, 45 years later: 
when you have dinner with the Knuths you talk more about quilting 
and printing Lutheran bibles than programming, 
but it all seems of a piece.}
\author GRAHAM NELSON (2018)
}

\medskip

{\quoteformat{$\ldots$ a function of royal blood in the family of entire functions, 
whose distinguished properties separate it 
from its bourgeois brethren}
\author Sir EDMUND WHITTAKER (1915)
}

\medskip
%\vfil \break

{\quoteformat{Aut inveniam viam 
aut faciam}
\author HANNIBAL (218 a.C.n.)
}

\medskip

{\quoteformat{
By avoiding detailed proofs and analysis, the book will appeal
to those who wish to enrich their problem-solving arsenal. 
The book is ideal for people who do not have a very deep 
academic background in mathematics, and yet wish to use 
mathematics for work or research.}
\author DANIEL COHEN-OR (20??)
}

\medskip

{\quoteformat{
We dedicate this book
To our fellow citizens
Who, for love of truth,
Take from their own wants
By taxes and gifts,
And now and then send forth
One of themselves
As dedicated servant,
To forward the search
In to the mysteries and marvelous simplicities
Of this strange and beautiful Universe,
Our home.}
\author CHARLES W. MISNER, KIP S. THORNE, JOHN ARCHIBALD WHEELER (19??)
}

\medskip

{\quoteformat{
For many years the theory of automata was developing rapidly
and solving problems that were ostensibly related to computers; 
but real programmers could not care less about those theorems 
because Turing machines were so different from real machines.}
\author DONALD E. KNUTH (1971)
}

\medskip

{\quoteformat{
If we make an unbiased examination of the accomplishments
made by mathematicians to the real world of computer programming, 
we are forced to conclude that, so far, 
the theory has actually done more harm than good.}
\author DONALD E. KNUTH (1971)
}

\medskip

{\quoteformat{
Always be the worst guy in every band you're in.}
\author PAT METHENY (19??)
}

\medskip

{\quoteformat{
The final algorithm of concern to us in this section is long division, 
in which we want to divide $(m+n)$-place integers by $n$-place integers.
Here the ordinary pencil-and-paper method involves a certain amount of 
guesswork and ingenuity on the part of the person doing the division; 
we must either eliminate this guesswork from the algorithm or 
develop some theory to explain it more carefully.}
\author DONALD E. KNUTH (19??)
}

\vfil \break

{\quoteformat{
If we were to wait 
for convergence proofs 
and error estimates 
for the new methods, 
most of the computers now in use in technology and industry 
would come grinding to a halt.}
\author ROBERT D. RICHTMYER, K. W. MORTON (1957)
}

\medskip


{\quoteformat{
On the other hand, it is part of the author's contention 
that mathematics books ought to mention the fact 
that a Taylor's series is often a very poor way 
to compute a function.}
\author GEORGE ELMER FORSYTHE (1970)
}

\medskip

{\quoteformat{
Everyone has a plan
'till they get punched in the mouth}
\author MIKE TYSON (19??)
}

\medskip

{\quoteformat{
Choosing the right precision for a problem
where the choice matters
requires significant understanding
of floating-point computation.
If you don't have that understanding, 
get advice, 
take the time to learn, 
or use double and hope for the best.}
\author BJARNE STROUSTRUP (2013)
}

\medskip

{\quoteformat{
I was saying the background does not satisfy
any mathematically or statistically simple models 
and those are the only kinds 
that we like to use in our papers 
with all the equations in them.}
\author AZRIEL ROSENFELD (1998)
}

\medskip

{\quoteformat{
If an Error--Analysis exists,
it may be obvious,
or it may be obscure.
If it exists,
it could cost more to find and compute
than $R$ is worth}
\author WILLIAM KAHAN (2016)
}

\medskip

{\quoteformat{
The 2011 Fukushima radioactive signal was
embedded in the shallow 2011 snow layers of the Mt. Ortles cores.
This result further suggests
that this event may be 
a new glaciological time horizon 
at a hemispheric scale, 
although only in the short term, 
due to the short half-life
of the radionuclides likely involved 
such as ${}^{131}I$.}
\author PAOLO GABRIELLI et al. (2016)
}

\medskip

{\quoteformat{
I used exponential running time algorithms 
when linear alternatives existed, 
because you can still raise venture capital for an app 
that only does half of what it should do 
before it crashes 
and wedges your entire phone.}
\author JAMES MICKENS (201?)
}

\medskip

{\quoteformat{
At any rate, 
when attacking a new problem 
it is often wise to find some fairly obvious procedure 
that works, 
and then 
try to improve upon it.}
\author DONALD KNUTH (2013)
}

\medskip

{\quoteformat{
The story goes 
that Google engineers 
had the idea to create 
a language designed for their 
large-scale applications 
and started designing Go 
while they were waiting 
for their massive C++ projects 
to compile.}
\author GUILLAUME LE STUM, ANTONIO SCARAMUZZINO (2019)
}

\medskip

{\quoteformat{
Ho perso quattro minuti 
per l'entusiasmo dei tifosi 
che pregherei di non lanciarmi pi\`u mazzi di fiori, 
come hanno fatto in vetta al Sestriere, 
provocandomi un incaglio al cambio, 
poich\'e erano fiori legati con fil di ferro 
e sono andati a incepparlo.}
\author GINO BARTALI (1949)
}

\medskip

{\quoteformat{
NASA engineers quipped 
that no spacecraft could fly 
until the mass of paper documenting its construction 
and test 
equalled that of the flight hardware.}
\author JOHN WALKER (2019)
}

\medskip

{\quoteformat{
I also got to spend a month in Japan 
designing the first electronic-shutter IC 
for a single-lens-reflex camera for Canon. 
This circuit took the logarithm of three inputs
-film speed, aperture, and light intensity-
summed the result, stored it while the mirror went up, 
and then took the antilogarithm to generate the shutter speed. 
It used about 20 transistors in total 
and was probably my most elegant design. 
Nowadays, 
this is done with about half a million transistors 
in a microcontroller-
how prosaic.}
\author DAVE FULLAGAR (2007)
}

\medskip

{\quoteformat{
It's just that I'm not worthy 
of the depths of TAOCP.
I've read the preface of Volume 1 three or four times, 
and I've tried to imagine how it would feel to complete 
the entire series. 
I would disconnect from the Internet for a few months 
and move alone to a cabin on a Wyoming mountaintop 
with a ream of paper, 
a couple boxes of pencils, 
TAOCP, 
a few supplementary math books, 
and a 90 day supply of Adderall.
I would take long hikes in the mountains, 
thinking deeply about the fundamental algorithms. 
No laptop, no wifi, no electricity -
these are all distractions. 
After sunset I'd read and work problem sets by candlelight, 
and my dreams would be joyrides 
through a universe of harmonic numbers, 
binomial coefficients, 
and nonlinear data structures.}
\author CARL TASHIAN (2016)
}

\medskip

{\quoteformat{
I gave a presentation at a conference 
when I was a PhD student, 
and during a break 
Knuth 
came up to me. 
My heart was pounding with excitement 
at the idea he was interested in my research. 
Then he opened his mouth and asked 
``Do you know where the toilets are?''
(in fact, his actually used the word ``john''
rather than ``toilet'').
Never before have I felt such disappointment, 
until I realised that out of all the people at the conference, 
he thought I was most likely to be the expert 
on something (albeit where the toilets are).}
\author SPRINGY (2019)
}

\medskip

{\quoteformat{
read until you get to math you don't understand, 
then skip it and find the next section 
you can understand}
\author NICK DROZD (2019)
}

\medskip

{\quoteformat{
I remember reading of a completion 
for a paper on resolution of singularities of surface; 
Castelnuovo and Enriques were in the committee. 
Beppo Levi presented his famous paper 
on the resolution of singularities for surfaces; 
Enriques asked him for a couple of examples 
and was convinced; 
Castelnuovo was not. 
The discussion got heated. 
Enriques exclaimed 
``I am ready to cut my head if this does not work''
and Castelnuovo replied 
``I don't think that would prove it either''.}
\author ANGELO VISTOLI (2010)
}

\medskip

{\quoteformat{
The three dots ``. . .'' are called an ellipsis. 
They suggest a pattern 
that is supposed to be obvious to the reader. 
But obviousness is in the mind of the beholder, 
and may not align with the intentions of the author.}
\author HARRY LEWIS, ‎RACHEL ZAX (2019)
}

\medskip

{\quoteformat{
I wish I could send a man from 
Fortnum and Mason's 
out to you 
with a chunk of mortadella.}
\author HENRY JAMES (18??)
}

\medskip

{\quoteformat{
Education is perceived as a cost 
more than an investment.
It is easy to estimate the cost of training people.
How do you estimate the value
of the bugs that you didn't have, 
thanks to those well-trained people?
We are in a situation where 
the cost is visible, 
and the benefit hidden. 
Unfortunately,
most people think that 
if they don't see the benefit, 
there is no benefit at all.}
\author J-P. ROSEN (2009)
}

\medskip

{\quoteformat{
Software development practices are designed
to mitigate the problems caused
by the inherent complexity of writing computer programs. 
Unfortunately, 
many methodologies veer off
into dogma,
bean counting, 
or both.}
\author GELMAN et al. (2020)
}



\bye